### 2016-01-31 article/main.tex

\documentclass[12pt,doc]{apa6}

\usepackage{blindtext}
\usepackage[american]{babel}
\usepackage[style=chicago,
            backend=biber,
            natbib=true,
            url=true]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\usepackage{txfonts}
\usepackage{microtype}

\bibliography{main.bib}
\title{Ethical Academic Software Practices: the Case for Behavior-Driven
Development}
\author{David Kraemer}
\date{November 24, 2015}
\shorttitle{Behavior-Driven Development in Academia}
\affiliation{Grinnell College}
\leftheader{Kraemer}
\doublespacing

\begin{document}

    \maketitle

       \printbibliography

\end{document}